Almost compatible functions and infinite length games

### 2015-11-18 articles/almost-compatible-functions-and-topological-games/article.tex
get kurepa family => S'(kappa)
writing up alan's proof of omega_n topology

    %TODO proof
    We now want to demonstrate that \(f_\alpha\sim f_{\beta_n}\) for all
    \(n<\omega\). We again need only concern ourselves with \(\delta_{m,j}\) for
    \(m\leq n\) since otherwise \(\delta_{m,j}\not\in K_{\beta_n}\).
    For \(m=n\), we have
    \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) which differs
    from \(f_{\beta_n}(\delta_{n,j})\) for only the finitely many \(j\) which
    are mapped below \(n\) by \(f_{\beta_n}\).
    For \(m<n\) and \(\delta_{m,j}\in K_{\beta_n}\),we have
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) which can
    only differ
    from \(f_{\beta_n}(\delta_{m,j})\) for only the finitely many \(j\) which
    are mapped below \(m\) by \(f_{\beta_m}\) or the finitely many \(j\)
    for which the
    almost compatible \(f_{\beta_n}\sim f_{\beta_m}\) differ.
  \end{proof}

@@

  % TODO: read through Alan's proof and fit to rest of document
          \begin{theorem}
           For each $n\in \omega$, there is a locally countable,
            $\omega$-bounded topology on $\omega_n$. Note
            that this means that the closure of any set has
            the same cardinality and weight as the set.
          \end{theorem}

          To prove the theorem, we must actually prove a stronger lemma.

          \begin{lemma}
          Assume that $X$ has cardinality at most $\omega_n$ (for
          any $n\in \omega$),
          and is locally countable, locally compact, and the closure
          of each set has the same cardinality as the set.
          Then $X$ has an $\omega$-bounded extension with the
          same properties.
          \end{lemma}

            \begin{proof}
            We prove this by induction on $n$.
            In fact, we make our inductive statement that  if $\tilde X$
            is the extension of $X$, then  $\tilde X \setminus X$ also
            has cardinality $\omega_n$.
             If $n=0$,
              then we can just take the free union of two copies of $X$
              and then the one-point compactification.
             So suppose $n>0$ and
            that $X$ is such a topology on the ordinal $\omega_n$.
            For each $\alpha<\omega_n$, the closure of the initial
            segment $\alpha$ is bounded by some $\gamma_\alpha$.
            Also, because $X$ is locally countable, $\gamma_\alpha$
            can be chosen so that $\alpha$ is contained  in the interior
            of $\gamma_\alpha$. There is a cub  $C\subset \omega_n$
            with the property that for each $\delta\in C$ and $\alpha<\delta$,
             $\gamma_\alpha$ is also less than $\delta$. This implies
             that for each $\delta\in C$, the initial segment $\delta$ is
             open, and if $\delta$ has uncountable cofinality, then
              $\delta$ is clopen.

              The proof will be easier to visualize if we now
              identify the points of $X$ with the point set
               $\omega_n\times \{0\}$ and we will add the points
                $\omega_n\times \{1\}$ to create the extension.
             By induction on $\lambda \in C$ we define a topology
             on $\omega_n\times\{0\}\cup \lambda\times \{1\}$ so that
             $\omega_n \times \{0\}$ is an open subset. We also ensure,
             by induction, for each $\alpha<\lambda$,
             the closure of $\alpha\times 2$ is an
              $\omega$-bounded  subset of $\lambda\times 2$.

          In the case that $n=1$, then  choose any sequence
           $\lambda_n : n\in \omega$ increasing cofinal in $\lambda$.
           If $\lambda $ is a limit in $C$, then we simply take the
           topology we have constructed so far on $\lambda\times 2$
           and there's nothing more needs to be done. Otherwise
          we may assume that $\lambda_0$ is the predecessor
          of $\lambda\in C$ and
           we set $Y_\lambda $ to equal the countable set
            $\overline{\lambda}\setminus \lambda$. For convenience,
             and with no loss, we assume that $\lambda$ itself is a limit
             of limits.
           And we have a topology on
           $$\lambda_0\times 2 \cup (\lambda \cup Y_\lambda)\times\{0\}~.$$
            Recursively
            choose clopen sets $U_n$ in this topology
             so that $\lambda_0\times 2\subset U_0$,
             $U_{n}\cup \lambda_{n+1}\times \{0\}$ is contained
              $ U_{n+1}$ while $U_{n+1}$ is disjoint from $Y_\lambda$.
              It is easy to see that we can have all the points
              in $(\lambda\setminus \{\lambda_n : n\in\omega\})\times \{1\}$
              be isolated, and arrange that $(\lambda_n,1) $ is the
              point at infinity in the one-point compactification
              $U_n\cup (\lambda_n\times \{1\})$.

                Now we handle the case $n>1$ and we can shrink $C$
                and now assume
                that $C$ is the closure of $\{\lambda\in C : \mbox{cf}(\lambda)>\omega\}$.
                We again proceed by induction on $\lambda\in C$.  If
                $\lambda$ is a limit in $C$, then there is nothing to do: we
                simply have defined an appropriate topology on
                $\omega_n\times  \{0\}\cup \lambda\times \{1\}$ so that
                for each $\mu\in C\cap \lambda$ with $\mbox{cf}(\mu)>\omega$,
                 $\mu\times 2$ is a clopen $\omega$-bounded subspace.
                In case $\lambda$ is not a limit of $C$, then $\lambda$
                has uncountable cofinality and a predecessor $\mu\in C$.
                We therefore have that $\lambda\times\{0\}$ is clopen
                in $\omega_n\times \{0\}$.
                 We apply the induction hypothesis
                to the space $\lambda\times \{0\} \cup \mu\times 2$ to choose
                the topology on $\lambda\times 2$.

>>>>

    We now want to demonstrate that \(f_\alpha\sim f_{\beta_n}\) for all
    \(n<\omega\). Note \(\delta_{m,j}\in K_{\beta_n}\) implies
    \(m\leq n\).
    For \(m=n\), we have
    \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) which differs
    from \(f_{\beta_n}(\delta_{n,j})\) for only the finitely many \(j\) which
    are mapped below \(n\) by \(f_{\beta_n}\).
    For \(m<n\) and \(\delta_{m,j}\in K_{\beta_n}\),we have
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) which can
    only differ
    from \(f_{\beta_n}(\delta_{m,j})\) for only the finitely many \(j\) which
    are mapped below \(m\) by \(f_{\beta_m}\) or the finitely many \(j\)
    for which the
    almost compatible \(f_{\beta_n}\sim f_{\beta_m}\) differ.

    Finally for any \(\beta<\alpha\), we may conclude \(f_\alpha\sim f_\beta\)
    since there is some \(\beta_n\) with
    \(K_\alpha\cap K_\beta=K_\alpha\cap K_{\beta_n}\),
    \(f_\alpha\sim f_{\beta_n}\), and \(f_{\beta_n}\sim f_\beta\).
  \end{proof}

  We now construct a topology on \(\omega_n\) for each \(n<\omega\) which
  will witness a Kurepa family on \(\aleph_n\).

\begin{lemma}
    Let \(X\) be a Hausdorff space of cardinality less than
    \(\aleph_\omega\) which is locally countable, locally compact, and the
    closure of each set has the same cardinality as the set.
    Then \(X\) has an \(\omega\)-bounded extension \(\tilde X\)
    with the same properties,
    such that \(\tilde X\setminus X\) has the same cardinality
    as \(X\).
  \end{lemma}

  \begin{proof}
    We prove this by induction on \(n\).
    If \(n=0\), then we can just use the one-point compactification of two
    copies of \(X\).
    So suppose \(n>0\) and
    that \(X=\{\alpha:\alpha<\omega_n\}\) has such a topology.
    For each \(\alpha<\omega_n\), \(\gamma_\alpha\) may be chosen such that
    both the closure of \(\alpha\) in \(X\) and a countable neighborhood of
    \(\alpha\) are subsets of \(\gamma_\alpha\). Note that
    the set
    \(C'=\{\lambda<\omega_n:\alpha<\lambda\Rightarrow\gamma_\alpha<\lambda\}\)
    is a cub subset of \(\omega_n\).
    It will be convenient later to use the cub \(C=C'\cap L\), where
    \(L\) contains the limits of limit ordinals in \(\omega_n\).
    Now for each \(\lambda\in C\), the initial segment \(\lambda\) is
    open as \(\alpha<\lambda\) belongs to the neighborhood
    \(\gamma_\alpha\subsetneq\lambda\). Also, if \(\lambda\) has uncountable
    cofinality, then for \(\beta\geq\lambda\) and any countable neighborhood
    \(U\) of \(\beta\), \(U\cap\lambda=U\cap\alpha\)
    for some \(\alpha<\lambda\); thus \(U\setminus\cl\alpha\)
    is a neighborhood of \(\beta\) missing \(\lambda\), showing that \(\lambda\)
    is clopen.

    Let \(\tilde X=\omega_n\times 2\).
    By induction on \(\lambda \in C\) we will define a topology
    for \(\omega_n\times\{0\}\cup \lambda\times \{1\}\) such that
    \begin{itemize}
      \item \(\omega_n \times \{0\}\) is an open copy of \(X\),
      \item the space is locally compact, and
      \item for each \(\alpha<\lambda\), the closure of \(\alpha\times 2\) is
        an \(\omega\)-bounded  subset of \(\lambda\times 2\).
    \end{itemize}

    In the case that \(n=1\), if \(\lambda\) is a limit in \(C\),
    then we simply give
    \(
      \omega_n\times\{0\}\cup\lambda\times\{1\}
    \) the topology induced by the union of the previously defined
    \(
      \omega_n\times\{0\}\cup\lambda'\times\{1\}
    \).
    Otherwise we choose an increasing sequence of limit
    ordinals \(\{\alpha_k : k\in \omega\}\) with limit \(\lambda\)
    such that \(\alpha_0\) is the predecessor of
    \(\lambda\) in \(C\) (or \(0\) if \(\lambda\) is the least element of
    \(C\)).

    %TODO finish edits
    The subspace \(\cl\lambda\times\{0\}\cup\alpha_0\times\{1\}\) of \(X\)
    is countable and locally compact; therefore it is
    metrizable and zero-dimensional.
    So we may choose sets \(U_k\) for \(k<\omega\) which are
    clopen in this topology and satisfy
    \(
      \cl{\alpha_k}\times\{0\}\cup\alpha_0\times\{1\}
        \subseteq
      U_n
        \subseteq
      \lambda\times\{0\}\cup\alpha_0\times\{1\}
    \).

    We need only describe a base for the points in
    \((\lambda\setminus\alpha_0)\times\{1\}\); we do so by letting
    \(\<\alpha,1\>\) be isolated when \(\alpha\not\in\{\alpha_k:k<\omega\}\),
    and letting \(\<\alpha_k,1\>\) be a compactifying point for
    \(U_n\cup(\alpha_k\times\{1\})\).

    This results in a locally compact space: for \(\alpha\geq\lambda\),
    \(\<\alpha,0\>\) has its usual compact neighborhoods within the copy
    \(\omega_n\times\{0\}\) of \(X\).

    Now we handle the case \(n>1\) and we can shrink \(C\)
    and now assume
    that \(C\) is the closure of \(\{\lambda\in C : \mbox{cf}(\lambda)>\omega\}\).
    We again proceed by induction on \(\lambda\in C\).  If
    \(\lambda\) is a limit in \(C\), then there is nothing to do: we
    simply have defined an appropriate topology on
    \(\omega_n\times  \{0\}\cup \lambda\times \{1\}\) so that
    for each \(\mu\in C\cap \lambda\) with \(\mbox{cf}(\mu)>\omega\),
     \(\mu\times 2\) is a clopen \(\omega\)-bounded subspace.
    In case \(\lambda\) is not a limit of \(C\), then \(\lambda\)
    has uncountable cofinality and a predecessor \(\mu\in C\).
    We therefore have that \(\lambda\times\{0\}\) is clopen
    in \(\omega_n\times \{0\}\).
     We apply the induction hypothesis
    to the space \(\lambda\times \{0\} \cup \mu\times 2\) to choose
    the topology on \(\lambda\times 2\).
  \end{proof}

  \begin{theorem}
    For each $n\in \omega$, there is a locally countable,
    $\omega$-bounded topology on $\omega_n$. Note
    that this means that the closure of any set has
    the same cardinality and weight as the set.
  \end{theorem}

  \begin{proof}
    Take a topological sum of \(\aleph_{n+1}\) copies of \(\omega_n\)
    with an appropriate topology (a convergent sequence for \(\omega\), the
    result of this theorem for larger cardinals),
    and apply the previous lemma.

    TODO: write an argument for the note.
  \end{proof}

@@
